Effects of Water Permeability on the Hydrodynamics of a Vesicle Under Flow and Confinement

### 2020-12-17 vesicle_code/docs/permeableVes/main.tex
yny

\affil[a]{Department of Scientific Computing, Florida State University}
\affil[b]{Department of Mathematical Sciences, New Jeresy Institute of
Technology}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Semi-Permeable Vesicle in Stenosis}
Cells experience stretching and release messengers when they pass through capillary vessels \cite{Wan2008_PNAS,ForsythWan2011_PNAS,Pak2015_PNAS}.
Many cells exhibit extraordinary flexibility as they go through narrow channels \cite{AuStoreyMoore2016_PNAS}, and adjustment of area to volume ratio may be a key factor for successful passages.
Here we quantify the effect of water permeability on a vesicle going through a stenosis (a long microfluidic channel with a cross section smaller than the cell radius).
Based on the findings in 
%We consider the effect of stenosis on a semi-permeable vesicle. Inspired
%by an experiment that measures the effects of cells on the dynamical
%pressure-drop variations along a micrometer-sized
\cite{abk-fai-sto2006}, we first examining the effect of
semi-permeability on the pressure fluctuation across a stenosed
geometry. In this experiment, a pressure drop is used to squeeze a
red blood cell (RBC) through a constriction with diameter $5\mu$m and
length $45\mu$m. We set up a similar computational geometry with a
Poiseuille flow imposed at the inlet and outlet
(Figure~\ref{fig:stenosisComposite}(a)). The flow rate is chosen so that
the residency time of the RBC in the constriction agrees with the
experimental results. When the RBC enters the constriction, the result
is an increase in the pressure drop required to maintain a constant flow
rate, and the difference between these pressures is called the excess
pressure. We consider both an impermeable and semi-permeable vesicle
with initial reduced area $\alpha = 0.65$ passing through a constricted
geometry and compute the pressure drop between the inlet and outlet of
the constriction. In Figure~\ref{fig:stenosisComposite}(c), we plot the
excess pressures. The impermeable case agrees with experimental
results~\cite{abk-fai-sto2006} (see Figure 3), and the results indicate
that less pressure is required to maintain a constant flow rate when the
vesicle is semi-permeable. Our simulations indicate the
semi-permeability does affect the excess pressure, and the vesicle's
reduced area reduces by about 6.1\%
(Figure~\ref{fig:stenosisComposite}(d)). Since the area does decrease at
a constant rate we expect that a larger amounts of deflation is possible
if the vesicle is stenosed for a longer period of time. Therefore, in
Figure~\ref{fig:stenosisComposite}(b), we consider a vesicle passing
through a much longer constriction, and we observe that the reduced area
decreases by 13.2\% (Figure~\ref{fig:stenosisComposite}(e)).

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Semi-Permeable Vesicle in a Contracting Geometry} 
We expose a single semi-permeable vesicle to even larger deformations by
simulating its hydrodynamics in a contracting microfluidic geometry.
Similar to~\cite{wu2015critical}, we consider a channel whose width
gradually reduces from $16\mu$m to $1.6\mu$m and then immediately opens
again to $16\mu$m (Figure~\ref{fig:contractingComposite}). We impose a
Poiseuille flow at the inlet and outlet with a constant maximum flow
rate of $U_{\max}$. We initialize a semi-permeable vesicle with $\alpha
= 0.65$ that spans $10\mu$m of the total channel width and investigate
the effects of the permeability rate and the maximum flow velocity. We
color the vesicle according to the sign of its flux---fluid flows into
the vesicle in red regions and out of the vesicle in blue regions.

In Figures~\ref{fig:contractingComposite}a
and~\ref{fig:contractingComposite}b, we plot snapshots of vesicles with
different permeability rates, and Figure~\ref{fig:contractingComposite}c
shows that the reduced area as a function of the vesicles' location. For
sufficiently large permeability rates, the vesicle initially inflates,
but then deflates as it passes through the constriction. However, if the
semi-permeability rate is too low, the vesicle is too large when it
reaches the narrowest part of the channel, and it is unable to pass. In
Figures~\ref{fig:contractingComposite}d
and~\ref{fig:contractingComposite}e, we plot snapshots of vesicles with
different imposed maximum velocities, and
Figure~\ref{fig:contractingComposite}f shows that the reduced area as a
function of the vesicles' location. At low velocities, the vesicle
inflates similar to the unbounded parabolic flow example, and it is
unable to pass through the constrictions. However, higher flow rates
result in additional deflation and the vesicle passes through the
contraction.

To understand long term effects of a vesicle passing through a
contracting geometry, in Figures~\ref{fig:contractingComposite}g we
simulate a vesicle passing six times through the contracting geometry.
In Figure~\ref{fig:contractingComposite}h, we show that the vesicle
deflates each time it passes through the constriction followed by a
rapid inflation. The total area loss over the history of the vesicle is
almost 50\%. In Figure~\ref{fig:contractingComposite}i, we show the
vesicle velocity as a function of the vesicle's center of mass. As the
vesicle passes through the contraction, it accelerates followed by a
rapid deacceleration once it has passed through the contraction.

>>>>

\affil[a]{Department of Scientific Computing, Florida State University}
\affil[b]{Department of Mathematical Sciences, New Jersey Institute of
Technology}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Semi-Permeable Vesicle in a Narrow Channel (Stenosis)}
Cells experience stretching and release messengers when they pass through capillary vessels \cite{Wan2008_PNAS,ForsythWan2011_PNAS,Pak2015_PNAS}.
Many cells exhibit extraordinary flexibility as they go through narrow channels \cite{AuStoreyMoore2016_PNAS}, and adjustment of area to volume ratio may be a key factor for a successful passage.
Here we quantify the effect of water permeability on a vesicle going through a stenosis (a long microfluidic channel with a width smaller than the cell radius).
Based on the findings in 
%We consider the effect of stenosis on a semi-permeable vesicle. Inspired
%by an experiment that measures the effects of cells on the dynamical
%pressure-drop variations along a micrometer-sized
\cite{abk-fai-sto2006}, we first examining the effect of
semi-permeability on the pressure fluctuation across a stenosed
geometry. In this experiment, a pressure drop is used to squeeze a
red blood cell (RBC) through a constriction with diameter $5\mu$m and
length $45\mu$m. We set up a similar computational geometry with a
Poiseuille flow imposed at the inlet and outlet
(Figure~\ref{fig:stenosisComposite}(a)). The flow rate is chosen so that
the residency time of the RBC in the constriction agrees with the
experimental results. When the RBC enters the channel, an increase in the pressure drop is required to maintain a constant flow
rate, and the difference in the pressure drop is called the excess
pressure.

We consider both an impermeable and semi-permeable vesicle
with initial reduced area $\alpha = 0.65$ passing through a constricted
geometry and compute the excess pressure as a function of vesicle position in the channel, see Figure~\ref{fig:stenosisComposite}(c). 
The impermeable case agrees with experimental
results in Figure 3 of \cite{abk-fai-sto2006}, and the results indicate
that less excess pressure is required to maintain a constant flow rate when the
vesicle is semi-permeable. Our simulations indicate the
semi-permeability does affect the excess pressure, and the vesicle's
reduced area reduces by about 6.1\%
(Figure~\ref{fig:stenosisComposite}(d)). Since the area does decrease at
a constant rate we expect that a larger amounts of deflation is possible
if the vesicle is stenosed for a longer period of time. Therefore, in
Figure~\ref{fig:stenosisComposite}(b), we consider a vesicle passing
through a channel of double the length, and we observe that the reduced area
decreases by 13.2\% (Figure~\ref{fig:stenosisComposite}(e)).

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Semi-Permeable Vesicle in a Contracting Geometry} 
Squeezing a cell or an elastic vesicle through a slit in microfluidics has gained popularity for its application in stress-induced release of macromolecules
\cite{ShareiEtAl2013_PNAS,Pak2015_PNAS,ZhangShenHoganBarakatMisbah2018_BJ,LuoBai2019_PoF}. RBCs go through submicron slits many times during their lifetime, and such process selects healthy, flexible RBCs to survive \cite{wu2015critical,LuPeng2019_PoF}.
Here we investigate the role of membrane permeability to water when a cell is squeezed through a contracting microfluidic channel similar to~\cite{wu2015critical},
where the width of the channel gradually decreases from $16\mu$m to $1.6\mu$m. The channel then immediately opens
again to $16\mu$m (Figure~\ref{fig:contractingComposite}). A
Poiseuille flow is imposed at the inlet and outlet with a constant maximum flow
rate of $U_{\max}$. We initialize a semi-permeable vesicle with $\alpha
= 0.65$ that spans $10\mu$m of the total channel width and investigate
the effects of the permeability rate and the maximum flow velocity. We
color the vesicle according to the sign of its flux---fluid flows into
the vesicle in red regions and out of the vesicle in blue regions.

In Figures~\ref{fig:contractingComposite}a
and~\ref{fig:contractingComposite}b, we plot snapshots of vesicles with
different permeability rates and membrane is color coded by the permeable flux (red for influx and blue for eflux).
Figure~\ref{fig:contractingComposite}c
shows the corresponding reduced area as a function of the vesicles' location. 
The reduced area in Figure~\ref{fig:contractingComposite}c and f shows that the vesicle always inflates
before it reaches the neck.  However, if the
semi-permeability rate is too low, the vesicle is too large when it
reaches the narrowest part of the channel, and it is unable to pass. In
Figures~\ref{fig:contractingComposite}d
and~\ref{fig:contractingComposite}e, we plot snapshots of vesicles with
different imposed maximum velocities, and
Figure~\ref{fig:contractingComposite}f shows that the reduced area as a
function of the vesicles' location. At low velocities, the vesicle
inflates similar to the unbounded parabolic flow example, and it is
unable to pass through the constrictions. However, higher flow rates
result in additional deflation and the vesicle passes through the
contraction.

To understand the effects of water permeability on cells that repeatedly go through micro-capillary vessels or sub-micron slits, 
we construct a contracting geometry (see Figures~\ref{fig:contractingComposite}g)
to simulate a vesicle passing through the contracting geometry six times.
In Figure~\ref{fig:contractingComposite}h, we show that the vesicle
deflates each time it passes through the constriction followed by a
rapid inflation. The total area loss over the history of the vesicle is
almost 50\%. In Figure~\ref{fig:contractingComposite}i, we show the
vesicle velocity as a function of the vesicle's center of mass. As the
vesicle passes through the contraction, it accelerates followed by a
rapid deacceleration once it has passed through the contraction.
